Wide Post-Mass Transfer White-Dwarf Binaries

### 2024-07-10 summaries/paper-draft-plain.tex
change paper draft stable mass section

\begin{tabular}{cl|cl}
  \verb|kstar| & evolutionary state & \verb|evol_type| & evolutionary change\\
  \hline
  0 & MS $<0.7 \Msun$ & 1 & initial state \\
  1 & MS $>0.7 \Msun$ & 2 & kstar change \\
  2 & Hertzprung Gap & 3 & begin Roche lobe overflow\\
  3 & First Giant Branch & 4 & end Roche lobe overflow\\
  4 & Core Helium Burning & 5 & contact \\
  5 & Early AGB & 6 & coalescence \\
  6 & TP-AGB & 7 & begin common envelope \\
  7 & Naked Helium Star MS & 8 & end common envelope \\
  8 & Naked Helium Star Hertzprung Gap & 9 & no remnant leftover \\
  9 & Naked Helium Star Giant Branch & 10 & max evolution time \\
  10 & Helium White Dwarf & 11 & binary disruption \\
  11 & Carbon/Oxygen White Dwarf & 12 & begin symbiotic phase \\
  12 & Oxygen/Neon White Dwarf & 13 & end symbiotic phase \\
  13 & Neutron Star & 14 & blue straggler \\
  14 & Black Hole & 15 & supernova of primary \\
  15 & Massless Remnant & 16 & supernova of secondary
\end{tabular}

@@

\subsubsection{Methods and results}
According to \cite{yamaguchi_lo}, the WD mass in the Gaia sample ranges from $0.5\Msun$ to $0.8\Msun$, which corresponds to progenitor mass of $1 \Msun$ to $3 \Msun$. The companions mass has a median of $0.85 \Msun$. MESA results for $1.5\Msun + 0.85\Msun$ systems is documented in \cite{yamaguchi_hi}. Hence, we consider a $1.5\Msun + 0.85\Msun$ evolution model in COSMIC and compare the results with those in \cite{yamaguchi_hi}. For the common envelope, we include all thermal energy $\Eth$ into the binding energy $\Ebind$, while the recombination energy $\Erec$ is included in part with a parameter $\alpharec$. That is,
\[
  \Ebind = \Egrav + \Eth + \alpharec \Erec.
\]

@@

After exploring the formation of WD+MS binaries through common envelope, we investigate the possibility of forming WD+MS binaries through stable mass transfer. In \cite{yamaguchi_lo}, the author plot the $\MWD - \Porb$ relation of observed wide WD+MS bianries, and argues that it is not possible to form wide WD+MS binaries through stable mass. The plot is recreated in Figure \ref{theory-observed}, and from the figure, we can see that the observed wide post-mass transfer WD+MS bianries do not match the theoretical $\MWD - \Porb$ relation for stable mass transfer calculated in \cite{rappaport1995relation}. Also, the author aruges that stable mass transfer requires donors on SGB at the onset of mass transfer. However, this will not lead to WDs with masses observed \cite{yamaguchi_lo}.

@@

As in the common envelope section, we investigate how the formation of WD+MS binaries depends on various parameters in high mass systems using $7\Msun + 1 \Msun$ binaries, and in low mass systems using $1.5\Msun + 0.85\Msun$ binaries.

>>>>

\begin{tabular}{cl | cl}
  \verb|kstar| & \textbf{evolutionary state} & \verb|evol_type| & \textbf{evolutionary change} \\
  \hline
  0 & MS $<0.7 \Msun$ & 1 & initial state \\
  1 & MS $>0.7 \Msun$ & 2 & kstar change \\
  2 & Hertzprung Gap & 3 & begin Roche lobe overflow\\
  3 & First Giant Branch & 4 & end Roche lobe overflow\\
  4 & Core Helium Burning & 5 & contact \\
  5 & Early AGB & 6 & coalescence \\
  6 & TP-AGB & 7 & begin common envelope \\
  7 & Naked Helium Star MS & 8 & end common envelope \\
  8 & Naked Helium Star Hertzprung Gap & 9 & no remnant leftover \\
  9 & Naked Helium Star Giant Branch & 10 & max evolution time \\
  10 & Helium White Dwarf & 11 & binary disruption \\
  11 & Carbon/Oxygen White Dwarf & 12 & begin symbiotic phase \\
  12 & Oxygen/Neon White Dwarf & 13 & end symbiotic phase \\
  13 & Neutron Star & 14 & blue straggler \\
  14 & Black Hole & 15 & supernova of primary \\
  15 & Massless Remnant & 16 & supernova of secondary
\end{tabular}

@@

\subsubsection{Methods and Results}
According to \cite{yamaguchi_lo}, the WD mass in the Gaia sample ranges from $0.5\Msun$ to $0.8\Msun$, which corresponds to progenitor mass of $1 \Msun$ to $3 \Msun$. The companions mass has a median of $0.85 \Msun$. MESA results for $1.5\Msun + 0.85\Msun$ systems is documented in \cite{yamaguchi_hi}. Hence, we consider a $1.5\Msun + 0.85\Msun$ evolution model in COSMIC and compare the results with those in \cite{yamaguchi_hi}. For the common envelope, we include all thermal energy $\Eth$ into the binding energy $\Ebind$, while the recombination energy $\Erec$ is included in part with a parameter $\alpharec$. That is,
\[
  \Ebind = \Egrav + \Eth + \alpharec \Erec.
\]

@@

After exploring the formation of WD+MS binaries through common envelope, we investigate the possibility of forming WD+MS binaries through stable mass transfer. In \cite{yamaguchi_lo}, the author plot the $\MWD - \Porb$ relation of observed wide WD+MS bianries, and argues that it is not possible to form wide WD+MS binaries through stable mass. The plot is recreated in Figure \ref{theory-observed}, and from the figure, we can see that the observed wide post-mass transfer WD+MS bianries do not match the theoretical $\MWD - \Porb$ relation for stable mass transfer calculated in \cite{rappaport1995relation}. Also, the author aruges that stable mass transfer requires donors on SGB at the onset of mass transfer. However, this will not lead to WDs with masses observed \cite{yamaguchi_lo}. We will verfiy this idea in this section using $7\Msun + 1 \Msun$ binaries.

@@

\subsection{Methods and Results}
First of all, to ensure stable mass transfer in COSMIC, we modify every entry of \verb|qcrit_array| in \verb|BSEDict| from the default $0.0$ to $20.0$. This changes the critical mass ratio for onset of unstable mass transfer to $20.0$ Based on our initial mass ratio $7:1$, this will make the actual mass raio of our system below this critical throughout the evolutionary process.

Secondly, a parameter on which stable mass transfer process depends is the accrection efficiency $\beta$, which in COSMIC is represented by \verb|acc_lim| in \verb|BSEDict|. Accretion efficiency is the fraction $\beta$ of all transfered materials that is accreted by the accretor. That is, $\dot{M_a} = -\beta\dot{M_d}$ where $M_d$ is the mass of donor and $M_a$ is the mass of accretor.

In our simulation, we investigate how accretion efficiency $\beta$ and initial separation $a_i$ affects the final separation $a_f$. We run a grid of three different $\beta$ — $\beta = 0.0$, $0.5$ and $1.0$, and a linear grid of $a_i$ from $2 \au$ to $8 \au$ with $400$ steps. In total, we simulate $1200$ binaries with different accretion limit and initial separation. For each system, after the simulation is finished, we select the first timestep, if any, that shows a WD+MS binary. We record the separation at this moment as the final separation $a_f$, and we compare them with the observed minimum separation $0.15 \au$ for post-mass transfer WD binaries.

The results are presented in Figure \ref{stable_hi}. Notice the gap for $a_i$ ranges in $3 \sim 4.5 \au$. This means no WD+MS forms for this initial separation. In fact, for these initial separations and accretion efficiencies, two star merges.

\begin{figure}
  \centering
  \includegraphics[width=0.75\linewidth]{fig/stable_hi_log.png}
  \caption{COSMIC results of WD+MS binaries after stable mass transfer for $7\Msun + 1\Msun$ systems, with accretion efficiency $\beta = 0$, $\beta = 0.5$, and $\beta = 1$ respectively. Left panel: dependence of final separation $a_f$ of the WD+MS binary on initial separation $a_i$. Right panel: The corresponding companion mass in solar units in the WD+MS binary. The gap in both figure represents that no WD+MS system forms for that initial separation.}
  \label{stable_hi}
\end{figure}

Notice that the majority of these systems have final separation $a_f > 0.15 \au$. However, this is \textbf{not the correct approach}, since the companion mass in our observed wide WD+MS binaries has median mass $\sim 1\Msun$. However, in the simulation results shown in Figure \ref{stable_hi}, the final companion mass is about $4 \Msun$ and $6.5 \Msun$ for $\beta = 0.5$ and $\beta = 1.0$ respectively. Hence, we need to adjust the  initial companion mass $M_{\star, i}$ and the accretion efficiency \verb|acc_lim|, in order to make the final companion mass $M_{\star, f}$ close to $1\Msun$. Therefore, we lower the initial companion mass $M_{\star, i}$ and adjust the accretion efficiency $\beta$ and re-run the simulation. The results for $7\Msun + 0.7\Msun$ with $\beta = 0.1$, and $7\Msun + 0.5\Msun$ with $\beta = 0.2$ is shown in Figure \ref{stable_hi_ex}. After adjusting the initial mass of companion and the accretion efficiency, now the final mass of the companion is about $1\Msun$ and we get systems similar to our observed sample.

\begin{figure}
  \centering
  \includegraphics[width = 0.75\linewidth]{fig/stable7+1ex.png}
  \caption{COSMIC results of WD+MS binaries after stable mass transfer for two systems — $7\Msun + 0.7\Msun$ with accretion efficiency $\beta = 0.1$, and $7\Msun + 0.5\Msun$ with accretion efficiency $\beta = 0.2$. The left and right panel presents the same content as in Figure \ref{stable_hi}.}
  \label{stable_hi_ex}
\end{figure}
